The Unicode Cookbook \linebreak for Linguists

### 2018-07-30 book/chapters/introduction.tex
minor edits in introduction

To also allow for different uppercase and lowercase letters and for a large
variety of control characters to be used in the newly developing technology of
computers, the American Standards Association decided to propose a new 7-bit
encoding in 1963 (with $2^7 = 128$ different possible characters), known as the
\textsc{American Standard Code for Information Interchange} (ASCII), geared
towards the encoding of English orthography. With the ascent of other
orthographies in computer usage, the wish to encode further variations of Latin
letters (including German <ß> and various letters with diacritics, e.g.\ <è>) led the
Digital Equipment Corporation to introduce an 8-bit \textsc{Multinational
Character Set} (MCS, with $2^8 = 256$ different possible characters), first used
with the introduction of the VT{\large 220} Terminal in 1983.

@@

In the 1980s various people started to develop true
international code sets. In the United States, a group of computer scientists
formed the \textsc{unicode consortium}, proposing a 16-bit encoding in 1991
(with $2^{16} = 65,536$ different possible characters). At the same time in
Europe, the \textsc{international organization for standardization} (ISO) was
working on ISO~10646 to replace the ISO/IEC~8859 standard. Their first draft of
the \textsc{universal character set} (UCS) in 1990 was 31-bit (with
theoretically $2^{31} = 2,147,483,648$ possible characters, but because of some
technical restrictions only 679,477,248 were allowed). Since 1991, the Unicode
Consortium and the ISO jointly develop the \textsc{unicode standard}, or
ISO/IEC~10646, leading to the current system including the original 16-bit
Unicode proposal as the \textsc{basic multilingual plane}, and 16 additional
planes of 16-bit for further extensions (with in total $(1+16) \cdot 2^{16} =
1,114,112$ possible characters). The most recent version of the Unicode Standard
(currently at version number 11.0.0) was published in June 2018 and it defines
137,374 different characters \citep{Unicode2018}.

>>>>

To also allow for different uppercase and lowercase letters and for a large
variety of control characters to be used in the newly developing technology of
computers, the American Standards Association decided to propose a new 7-bit
encoding in 1963 (with 2\textsuperscript{7} = 128 different possible characters), 
known as the
\textsc{American Standard Code for Information Interchange} (ASCII), geared
towards the encoding of English orthography. With the ascent of other
orthographies in computer usage, the wish to encode further variations of Latin
letters (including German <ß> and various letters with diacritics, e.g.\ <è>) led the
Digital Equipment Corporation to introduce an 8-bit \textsc{Multinational
Character Set} (MCS, with 2\textsuperscript{8} = 256 different possible characters), first used with the introduction of the VT{\large 220} Terminal in 1983.

@@

In the 1980s various people started to develop true
international code sets. In the United States, a group of computer scientists
formed the \textsc{unicode consortium}, proposing a 16-bit encoding in 1991
(with 2\textsuperscript{16} = 65,536 different possible characters). At the same time in
Europe, the \textsc{international organization for standardization} (ISO) was
working on ISO~10646 to replace the ISO/IEC~8859 standard. Their first draft of
the \textsc{universal character set} (UCS) in 1990 was 31-bit (with
theoretically 2\textsuperscript{31} = 2,147,483,648 possible characters, but because of some
technical restrictions only 679,477,248 were allowed). Since 1991, the Unicode
Consortium and the ISO jointly develop the \textsc{unicode standard}, or
ISO/IEC~10646, leading to the current system including the original 16-bit
Unicode proposal as the \textsc{basic multilingual plane}, and 16 additional
planes of 16-bit for further extensions (with in total (1+16)\times 2\textsuperscript{16} =
1,114,112 possible characters). The most recent version of the Unicode Standard
(currently at version number 11.0.0) was published in June 2018 and it defines
137,374 different characters \citep{Unicode2018}.
